### 2015-10-25 ch02-qlogms/gridDefs.inc.tex
Import QLogMS TIM paper

\definecolor{model}{named}{blue}
\definecolor{truesys}{named}{TangoAluminium5}

\definecolor{linms}{named}{TangoScarletRed3}
\definecolor{logms}{named}{TangoChameleon2}
\definecolor{qlogms}{named}{TangoPlum3}
\definecolor{clogms}{named}{TangoSkyBlue3}
\definecolor{plogms}{named}{black}

\definecolor{sigmalinms}{named}{linms}
\definecolor{sigmaqlogms}{named}{qlogms}
\definecolor{sigmaclogms}{named}{clogms}

\definecolor{biaslinms}{named}{TangoScarletRed1}
\definecolor{biasqlogms}{named}{TangoPlum1}
\definecolor{biasclogms}{named}{TangoSkyBlue1}

\definecolor{dbacolor}{named}{TangoScarletRed2}
\definecolor{dbbcolor}{named}{TangoSkyBlue2}
\definecolor{dbccolor}{named}{TangoChameleon2}
\definecolor{dbdcolor}{named}{TangoOrange2}

\pgfplotsset{linmsgrid/.style={ycomb,color=linms,solid,line width=0.7pt,mark size=1pt,mark=square*,mark options={solid}}}
\pgfplotsset{logmsgrid/.style={ycomb,color=logms,solid,line width=0.7pt,mark size=1pt,mark=*,mark options={solid}}}
\pgfplotsset{clogmsgrid/.style={ycomb,color=clogms,solid,line width=0.7pt,mark size=2pt,mark=x,mark options={solid}}}
\pgfplotsset{plogmsgrid/.style={ycomb,color=plogms,solid,line width=0.7pt,mark size=2pt,mark=triangle,mark options={solid,line width=0.7pt}}}

\pgfplotsset{model/.style={color=model,dashed,line width=1.0pt}}
\pgfplotsset{truesys/.style={color=truesys,line width=1.0pt}}

\newcommand{\linmsmarker}{mark=square,mark options={solid}}
\newcommand{\qlogmsmarker}{mark=triangle,mark options={solid}}
\newcommand{\clogmsmarker}{mark=x,mark size=10pt,mark options={solid}}

\newcommand{\dbamarker}{mark=x       ,mark options={solid}}
\newcommand{\dbbmarker}{mark=o       ,mark options={solid}}
\newcommand{\dbcmarker}{mark=square  ,mark options={solid}}
\newcommand{\dbdmarker}{mark=triangle,mark options={solid}}

\pgfplotsset{sigmalinms/.append style={color=sigmalinms,solid,line width=0.8pt}}
\pgfplotsset{sigmaqlogms/.append style={color=sigmaqlogms,solid,line width=0.8pt}}
\pgfplotsset{sigmaclogms/.append style={color=sigmaclogms,solid,line width=0.8pt}}

\pgfplotsset{biaslinms/.style={color=biaslinms,only marks,mark=x,mark options={solid}}}
\pgfplotsset{biasqlogms/.style={color=biasqlogms,only marks,mark=triangle,mark options={solid,rotate=180}}}
\pgfplotsset{biasclogms/.style={color=biasclogms,only marks,mark=o,mark options={solid}}}

\pgfplotsset{linms/.style={color=linms,solid,line width=0.5pt}}
\pgfplotsset{qlogms/.style={color=qlogms,solid,line width=0.5pt}}
\pgfplotsset{clogms/.style={color=clogms,solid,line width=0.5pt}}

\pgfplotsset{linmsconnect/.style={color=linms,%
                                  solid,
                                  line width=1pt,
                                  mark=square*,
                                  mark size=1pt,
                                  mark options={solid}}}
\pgfplotsset{qlogmsconnect/.style={color=qlogms,%
                                  solid,
                                  line width=1pt,%
                                  mark=*,%
                                  mark size=1pt,
                                  mark options={solid}}}
\pgfplotsset{clogmsconnect/.style={color=clogms,%
                                   solid,
                                   line width=1pt,%
                                   mark=x,%
                                   mark size=1pt,
                                   mark options={solid}}}

\pgfplotsset{dbavar/.style={color=dbacolor,solid,line width=1.0pt}}
\pgfplotsset{dbbvar/.style={color=dbbcolor,solid,line width=1.0pt}}
\pgfplotsset{dbcvar/.style={color=dbccolor,solid,line width=1.0pt}}
\pgfplotsset{dbdvar/.style={color=dbdcolor,solid,line width=1.0pt}}

\pgfplotsset{dbamin/.style={color=dbacolor,dashed,line width=1.0pt}}
\pgfplotsset{dbbmin/.style={color=dbbcolor,dashed,line width=1.0pt}}
\pgfplotsset{dbcmin/.style={color=dbccolor,dashed,line width=1.0pt}}
\pgfplotsset{dbdmin/.style={color=dbdcolor,dashed,line width=1.0pt}}

\pgfplotsset{dbamax/.style={color=dbacolor,solid,line width=1.0pt,mark=x ,mark options={solid}}}
\pgfplotsset{dbbmax/.style={color=dbbcolor,solid,line width=1.0pt,mark=o ,mark options={solid}}}
\pgfplotsset{dbcmax/.style={color=dbccolor,solid,line width=1.0pt,mark=square  ,mark options={solid}}}
\pgfplotsset{dbdmax/.style={color=dbdcolor,solid,line width=1.0pt,mark=triangle,mark options={solid}}}