### 2015-10-25 ch02-qlogms/grids.inc.tex

%% Colors
  \definecolor{tangored}{named}{TangoScarletRed1}
  \definecolor{tangoblue}{rgb}{0.203922,0.396078,0.643137}
  \definecolor{tangogreen}{named}{TangoChameleon2}
  \definecolor{tangogrey}{named}{TangoAluminium5}

  \definecolor{meas}{named}{TangoPlum2}

  \definecolor{goallight}{named}{TangoChameleon1}
  \definecolor{goal}{named}{TangoChameleon2}
  \definecolor{goaldark}{named}{TangoChameleon3}

  \definecolor{obtainedlight}{named}{TangoScarletRed1}
  \definecolor{obtained}{named}{TangoScarletRed2}
  \definecolor{obtaineddark}{named}{TangoScarletRed3}

  \definecolor{model}{named}{tangogrey}
  \definecolor{bg}{named}{TangoAluminium2}

\definecolor{linms}{named}{TangoScarletRed3}
\definecolor{logms}{named}{TangoChameleon2}
\definecolor{qlogms}{named}{TangoPlum3}
\definecolor{clogms}{named}{TangoSkyBlue3}
\definecolor{plogms}{named}{black}

  \definecolor{compen}{rgb}{0.729,0.741,0.714}
  \definecolor{signal}{named}{TangoSkyBlue2}
  \definecolor{misalign}{named}{TangoScarletRed2}

  \definecolor{dbacolor}{named}{TangoScarletRed2}
  \definecolor{dbbcolor}{named}{TangoSkyBlue2}
  \definecolor{dbccolor}{named}{TangoChameleon2}
  \definecolor{dbdcolor}{named}{TangoOrange2}

%% TikZ styles
  \tikzset{obtainedarrow/.style={color=obtaineddark,line width=1.0pt,->}}
  \tikzset{goalarrow/.style={color=goaldark,line width=1.0pt,->}}
  \tikzset{arrow/.style={color=TangoAluminium5,line width=1pt,->,>={triangle 45}}}

%% marker styles
  \definecolor{bwfill}{named}{TangoAluminium1}
  \definecolor{bwline}{named}{TangoAluminium5}

  \definecolor{safezone}{named}{bwfill}

%% pgfplots styles
\pgfplotsset{goal/.style={color=goallight,dashed,line width=0.75pt}}
\pgfplotsset{obtained/.style={color=obtainedlight,line width=0.75pt}}

\pgfplotsset{signalrectangle/.style={fill=signal!60!black,fill opacity=0.8,color=signal!30}}
\pgfplotsset{signalrectanglerobust/.style={fill=signal!50,fill opacity=0.5,color=signal!30}}

\pgfplotsset{safezone/.style={fill=safezone,fill opacity=0.5,color=safezone}}

\tikzset{showdistance/.style={|-|,line width=1pt}}

\pgfplotsset{frf/.style={color=black,solid,line width=1pt}}
\pgfplotsset{bw/.style={fill=bwfill,draw=bwline}}
\pgfplotsset{signal/.style={ycomb,color=signal,solid,line width=0.75pt}}

\pgfplotsset{dbvariance/.style={solid,line width=1.0pt}}
\pgfplotsset{dbvariancemin/.style={dashed,line width=1.0pt}}
\pgfplotsset{dbvariancemax/.style={dbvariance}}

\pgfplotsset{dbavar/.style={color=dbacolor,dbvariance}}
\pgfplotsset{dbbvar/.style={color=dbbcolor,dbvariance}}
\pgfplotsset{dbcvar/.style={color=dbccolor,dbvariance}}
\pgfplotsset{dbdvar/.style={color=dbdcolor,dbvariance}}

\pgfplotsset{dbamin/.style={color=dbacolor,dbvariancemin}}
\pgfplotsset{dbbmin/.style={color=dbbcolor,dbvariancemin}}
\pgfplotsset{dbcmin/.style={color=dbccolor,dbvariancemin}}
\pgfplotsset{dbdmin/.style={color=dbdcolor,dbvariancemin}}

\pgfplotsset{dbamax/.style={color=dbacolor,dbvariancemax,mark=diamond*}}
\pgfplotsset{dbbmax/.style={color=dbbcolor,dbvariancemax,mark=*}}
\pgfplotsset{dbcmax/.style={color=dbccolor,dbvariancemax,mark=square*}}
\pgfplotsset{dbdmax/.style={color=dbdcolor,dbvariancemax,mark=triangle*}}

\pgfplotsset{linmsmarker/.style={mark=*,mark options={solid},mark size=1pt}}
\pgfplotsset{logmsmarker/.style={mark=pentagon*,mark options={solid},mark size=1pt}}
\pgfplotsset{clogmsmarker/.style={mark=triangle*,mark options={solid},mark size=1pt}}
\pgfplotsset{qlogmsmarker/.style={mark=square*,mark options={solid},mark size=1pt}}
\pgfplotsset{plogmsmarker/.style={mark=diamond*,mark options={solid},mark size=1pt}}

\pgfplotsset{model/.style={color=model,line width=1.0pt}}

\pgfplotsset{gridline/.style={ycomb,solid,line width=0.7pt,mark size=1pt}}
\pgfplotsset{biggrid/.style={ycomb,solid,line width=1pt,mark size=1pt}}

\pgfplotsset{linmsgrid/.style={color=linms,gridline,mark=square*,mark options={solid}}}
\pgfplotsset{logmsgrid/.style={color=logms,gridline,mark=*,mark options={solid}}}
\pgfplotsset{clogmsgrid/.style={color=clogms,gridline,mark size=2pt,mark=x,mark options={solid}}}
\pgfplotsset{plogmsgrid/.style={color=plogms,gridline,mark size=2pt,mark=triangle,mark options={solid,line width=0.7pt}}}

\pgfplotsset{linms/.style={color=linms}}
\pgfplotsset{clogms/.style={color=clogms}}
\pgfplotsset{rlogms/.style={color=rlogms}}
\pgfplotsset{qlogms/.style={color=qlogms}}
\pgfplotsset{plogms/.style={color=plogms}}

\pgfplotsset{compen/.style={fill,fill opacity=0.65,color=compen,smooth,forget plot}}
\pgfplotsset{truesys/.style={color=tangogrey,line width=0.9pt}}

\pgfplotsset{meas/.style={color=meas,only marks,mark=*,mark options={solid},mark size=1pt}}

\pgfplotsset{results/.style={solid,line width=0.7pt}}
\pgfplotsset{linmsres/.style={color=linms,results}}
\pgfplotsset{qlogmsres/.style={color=qlogms,results}}
\pgfplotsset{clogmsres/.style={color=clogms,results}}
\pgfplotsset{plogmsres/.style={color=plogms,results}}

\pgfplotsset{linmsmeas/.style={color=linms,densely dashed,line width=0.7pt}}
\pgfplotsset{qlogmsmeas/.style={color=qlogms,densely dotted,line width=0.7pt}}
\pgfplotsset{clogmsmeas/.style={color=clogms,solid,line width=0.7pt}}
\pgfplotsset{plogmsmeas/.style={color=plogms,solid,line width=0.7pt}}

\pgfplotsset{connectline/.style={solid,line width=1.0pt}}
\pgfplotsset{linmsconnect/.style={color=linms,connectline,linmsmarker}}
\pgfplotsset{qlogmsconnect/.style={color=qlogms,connectline,qlogmsmarker}}
\pgfplotsset{clogmsconnect/.style={color=clogms,connectline,clogmsmarker}}
\pgfplotsset{plogmsconnect/.style={color=plogms,connectline,plogmsmarker}}

\pgfplotsset{sigmalinms/.style={linms,connectline}}
\pgfplotsset{sigmaqlogms/.style={qlogms,connectline}}
\pgfplotsset{sigmaclogms/.style={clogms,connectline}}
\pgfplotsset{sigmaplogms/.style={plogms,connectline}}

\pgfplotsset{linrect/.style={fill=linms,draw=linms,line width=1pt}}