### 2015-10-28 front-and-back-matter/titlepages.tex
Add title page, some bib changes

%\promotortitle{Promotor/Promotors}
\promotor{
  \jurymember{prof. dr. ir.}{Johan Schoukens}{Vrije Universiteit Brussel}
  \jurymember{dr. ir.}{Tom Oomen}{Eindhoven University of Technology}
}

\title{Development of User-friendly\\System Identification Techniques}

\author{ir. Egon Geerardyn}

\jury{
  \jurymember{prof. dr. ir.}{Keith R. Godfrey}{Warwick University}
  \jurymember{prof. dr. ir.}{Tomas McKelvey}{Chalmers}
  \jurymember{dr. ir.}{Bart Peeters}{LMS}
  \jurymember[president]{prof. dr. ir.}{Johan Deconinck}{Vrije Universiteit Brussel}
  \jurymember[secretary]{dr. ir.}{Koen Tiels}{Vrije Universiteit Brussel}
}

\date{\today~~{\color{gray}(\texttt{\VCRevision})}}

\maketitlepage