### 2015-10-28 config/styles.inc.tex

\pgfplotsset{compat=newest}

\def\thisDir{.}

\pgfplotsset{tick style={black!30},grid style={black!10}}
\pgfplotsset{every axis legend/.append style={font=\footnotesize}}
\pgfplotsset{every axis label/.append style={font=\footnotesize}}
\pgfplotsset{every tick label/.append style={font=\footnotesize}}
\pgfplotsset{every axis title/.append style={font=\footnotesize}}
\pgfplotsset{every axis post/.style={unbounded coords=jump}}
\pgfplotsset{every axis title/.append style={at={(0.5,0.95)}}}

%% Colors
  \definecolor{tangored}{named}{TangoScarletRed1}
  \definecolor{tangoblue}{rgb}{0.203922,0.396078,0.643137}
  \definecolor{tangogreen}{named}{TangoChameleon2}
  \definecolor{tangogrey}{named}{TangoAluminium5}

  \definecolor{meas}{named}{TangoPlum2}

  \definecolor{goallight}{named}{TangoChameleon1}
  \definecolor{goal}{named}{TangoChameleon2}
  \definecolor{goaldark}{named}{TangoChameleon3}

  \definecolor{obtainedlight}{named}{TangoScarletRed1}
  \definecolor{obtained}{named}{TangoScarletRed2}
  \definecolor{obtaineddark}{named}{TangoScarletRed3}

  \definecolor{model}{named}{tangogrey}
  \definecolor{bg}{named}{TangoAluminium2}

\definecolor{linms}{named}{TangoScarletRed3}
\definecolor{logms}{named}{TangoChameleon2}
\definecolor{qlogms}{named}{TangoPlum3}
\definecolor{clogms}{named}{TangoSkyBlue3}
\definecolor{plogms}{named}{black}

  \definecolor{compen}{rgb}{0.729,0.741,0.714}
  \definecolor{signal}{named}{TangoSkyBlue2}
  \definecolor{misalign}{named}{TangoScarletRed2}

  \definecolor{dbacolor}{named}{TangoScarletRed2}
  \definecolor{dbbcolor}{named}{TangoSkyBlue2}
  \definecolor{dbccolor}{named}{TangoChameleon2}
  \definecolor{dbdcolor}{named}{TangoOrange2}

%% TikZ styles
  \tikzset{obtainedarrow/.style={color=obtaineddark,line width=1.0pt,->}}
  \tikzset{goalarrow/.style={color=goaldark,line width=1.0pt,->}}
  \tikzset{arrow/.style={color=TangoAluminium5,line width=1pt,->,>={triangle 45}}}

%% marker styles
  \definecolor{bwfill}{named}{TangoAluminium1}
  \definecolor{bwline}{named}{TangoAluminium5}

  \definecolor{safezone}{named}{bwfill}

%% pgfplots styles
\pgfplotsset{goal/.style={color=goallight,dashed,line width=0.75pt}}
\pgfplotsset{obtained/.style={color=obtainedlight,line width=0.75pt}}

\pgfplotsset{signalrectangle/.style={fill=signal!60!black,fill opacity=0.8,color=signal!30}}
\pgfplotsset{signalrectanglerobust/.style={fill=signal!50,fill opacity=0.5,color=signal!30}}

\pgfplotsset{safezone/.style={fill=safezone,fill opacity=0.5,color=safezone}}

\tikzset{showdistance/.style={|-|,line width=1pt}}

\pgfplotsset{frf/.style={color=black,solid,line width=1pt}}
\pgfplotsset{bw/.style={fill=bwfill,draw=bwline}}
\pgfplotsset{signal/.style={ycomb,color=signal,solid,line width=0.75pt}}

\pgfplotsset{dbvariance/.style={solid,line width=1.0pt}}
\pgfplotsset{dbvariancemin/.style={dashed,line width=1.0pt}}
\pgfplotsset{dbvariancemax/.style={dbvariance}}

\pgfplotsset{dbavar/.style={color=dbacolor,dbvariance}}
\pgfplotsset{dbbvar/.style={color=dbbcolor,dbvariance}}
\pgfplotsset{dbcvar/.style={color=dbccolor,dbvariance}}
\pgfplotsset{dbdvar/.style={color=dbdcolor,dbvariance}}

\pgfplotsset{dbamin/.style={color=dbacolor,dbvariancemin}}
\pgfplotsset{dbbmin/.style={color=dbbcolor,dbvariancemin}}
\pgfplotsset{dbcmin/.style={color=dbccolor,dbvariancemin}}
\pgfplotsset{dbdmin/.style={color=dbdcolor,dbvariancemin}}

\pgfplotsset{dbamax/.style={color=dbacolor,dbvariancemax,mark=diamond*}}
\pgfplotsset{dbbmax/.style={color=dbbcolor,dbvariancemax,mark=*}}
\pgfplotsset{dbcmax/.style={color=dbccolor,dbvariancemax,mark=square*}}
\pgfplotsset{dbdmax/.style={color=dbdcolor,dbvariancemax,mark=triangle*}}

\pgfplotsset{linmsmarker/.style={mark=*,mark options={solid},mark size=1pt}}
\pgfplotsset{logmsmarker/.style={mark=pentagon*,mark options={solid},mark size=1pt}}
\pgfplotsset{clogmsmarker/.style={mark=triangle*,mark options={solid},mark size=1pt}}
\pgfplotsset{qlogmsmarker/.style={mark=square*,mark options={solid},mark size=1pt}}
\pgfplotsset{plogmsmarker/.style={mark=diamond*,mark options={solid},mark size=1pt}}

\pgfplotsset{model/.style={color=model,line width=1.0pt}}

\pgfplotsset{gridline/.style={ycomb,solid,line width=0.7pt,mark size=1pt}}
\pgfplotsset{biggrid/.style={ycomb,solid,line width=1pt,mark size=1pt}}

\pgfplotsset{linmsgrid/.style={color=linms,gridline,mark=square*,mark options={solid}}}
\pgfplotsset{logmsgrid/.style={color=logms,gridline,mark=*,mark options={solid}}}
\pgfplotsset{clogmsgrid/.style={color=clogms,gridline,mark size=2pt,mark=x,mark options={solid}}}
\pgfplotsset{plogmsgrid/.style={color=plogms,gridline,mark size=2pt,mark=triangle,mark options={solid,line width=0.7pt}}}

\pgfplotsset{linms/.style={color=linms}}
\pgfplotsset{clogms/.style={color=clogms}}
\pgfplotsset{rlogms/.style={color=rlogms}}
\pgfplotsset{qlogms/.style={color=qlogms}}
\pgfplotsset{plogms/.style={color=plogms}}

\pgfplotsset{compen/.style={fill,fill opacity=0.65,color=compen,smooth,forget plot}}
\pgfplotsset{truesys/.style={color=tangogrey,line width=0.9pt}}

\pgfplotsset{meas/.style={color=meas,only marks,mark=*,mark options={solid},mark size=1pt}}

\pgfplotsset{results/.style={solid,line width=0.7pt}}
\pgfplotsset{linmsres/.style={color=linms,results}}
\pgfplotsset{qlogmsres/.style={color=qlogms,results}}
\pgfplotsset{clogmsres/.style={color=clogms,results}}
\pgfplotsset{plogmsres/.style={color=plogms,results}}

\pgfplotsset{linmsmeas/.style={color=linms,densely dashed,line width=0.7pt}}
\pgfplotsset{qlogmsmeas/.style={color=qlogms,densely dotted,line width=0.7pt}}
\pgfplotsset{clogmsmeas/.style={color=clogms,solid,line width=0.7pt}}
\pgfplotsset{plogmsmeas/.style={color=plogms,solid,line width=0.7pt}}

\pgfplotsset{connectline/.style={solid,line width=1.0pt}}
\pgfplotsset{linmsconnect/.style={color=linms,connectline,linmsmarker}}
\pgfplotsset{qlogmsconnect/.style={color=qlogms,connectline,qlogmsmarker}}
\pgfplotsset{clogmsconnect/.style={color=clogms,connectline,clogmsmarker}}
\pgfplotsset{plogmsconnect/.style={color=plogms,connectline,plogmsmarker}}

\pgfplotsset{sigmalinms/.style={linms,connectline}}
\pgfplotsset{sigmaqlogms/.style={qlogms,connectline}}
\pgfplotsset{sigmaclogms/.style={clogms,connectline}}
\pgfplotsset{sigmaplogms/.style={plogms,connectline}}

\pgfplotsset{linrect/.style={fill=linms,draw=linms,line width=1pt}}

\definecolor{model}{named}{blue}
\definecolor{truesys}{named}{TangoAluminium5}

\definecolor{linms}{named}{TangoScarletRed3}
\definecolor{logms}{named}{TangoChameleon2}
\definecolor{qlogms}{named}{TangoPlum3}
\definecolor{clogms}{named}{TangoSkyBlue3}
\definecolor{plogms}{named}{black}

\definecolor{sigmalinms}{named}{linms}
\definecolor{sigmaqlogms}{named}{qlogms}
\definecolor{sigmaclogms}{named}{clogms}

\definecolor{biaslinms}{named}{TangoScarletRed1}
\definecolor{biasqlogms}{named}{TangoPlum1}
\definecolor{biasclogms}{named}{TangoSkyBlue1}

\definecolor{dbacolor}{named}{TangoScarletRed2}
\definecolor{dbbcolor}{named}{TangoSkyBlue2}
\definecolor{dbccolor}{named}{TangoChameleon2}
\definecolor{dbdcolor}{named}{TangoOrange2}

\pgfplotsset{linmsgrid/.style={ycomb,color=linms,solid,line width=0.7pt,mark size=1pt,mark=square*,mark options={solid}}}
\pgfplotsset{logmsgrid/.style={ycomb,color=logms,solid,line width=0.7pt,mark size=1pt,mark=*,mark options={solid}}}
\pgfplotsset{clogmsgrid/.style={ycomb,color=clogms,solid,line width=0.7pt,mark size=2pt,mark=x,mark options={solid}}}
\pgfplotsset{plogmsgrid/.style={ycomb,color=plogms,solid,line width=0.7pt,mark size=2pt,mark=triangle,mark options={solid,line width=0.7pt}}}

\pgfplotsset{model/.style={color=model,dashed,line width=1.0pt}}
\pgfplotsset{truesys/.style={color=truesys,line width=1.0pt}}

\newcommand{\linmsmarker}{mark=square,mark options={solid}}
\newcommand{\qlogmsmarker}{mark=triangle,mark options={solid}}
\newcommand{\clogmsmarker}{mark=x,mark size=10pt,mark options={solid}}

\newcommand{\dbamarker}{mark=x       ,mark options={solid}}
\newcommand{\dbbmarker}{mark=o       ,mark options={solid}}
\newcommand{\dbcmarker}{mark=square  ,mark options={solid}}
\newcommand{\dbdmarker}{mark=triangle,mark options={solid}}

\pgfplotsset{sigmalinms/.append style={color=sigmalinms,solid,line width=0.8pt}}
\pgfplotsset{sigmaqlogms/.append style={color=sigmaqlogms,solid,line width=0.8pt}}
\pgfplotsset{sigmaclogms/.append style={color=sigmaclogms,solid,line width=0.8pt}}

\pgfplotsset{biaslinms/.style={color=biaslinms,only marks,mark=x,mark options={solid}}}
\pgfplotsset{biasqlogms/.style={color=biasqlogms,only marks,mark=triangle,mark options={solid,rotate=180}}}
\pgfplotsset{biasclogms/.style={color=biasclogms,only marks,mark=o,mark options={solid}}}

\pgfplotsset{linms/.style={color=linms,solid,line width=0.5pt}}
\pgfplotsset{qlogms/.style={color=qlogms,solid,line width=0.5pt}}
\pgfplotsset{clogms/.style={color=clogms,solid,line width=0.5pt}}

\pgfplotsset{linmsconnect/.style={color=linms,%
                                  solid,
                                  line width=1pt,
                                  mark=square*,
                                  mark size=1pt,
                                  mark options={solid}}}
\pgfplotsset{qlogmsconnect/.style={color=qlogms,%
                                  solid,
                                  line width=1pt,%
                                  mark=*,%
                                  mark size=1pt,
                                  mark options={solid}}}
\pgfplotsset{clogmsconnect/.style={color=clogms,%
                                   solid,
                                   line width=1pt,%
                                   mark=x,%
                                   mark size=1pt,
                                   mark options={solid}}}

\pgfplotsset{dbavar/.style={color=dbacolor,solid,line width=1.0pt}}
\pgfplotsset{dbbvar/.style={color=dbbcolor,solid,line width=1.0pt}}
\pgfplotsset{dbcvar/.style={color=dbccolor,solid,line width=1.0pt}}
\pgfplotsset{dbdvar/.style={color=dbdcolor,solid,line width=1.0pt}}

\pgfplotsset{dbamin/.style={color=dbacolor,dashed,line width=1.0pt}}
\pgfplotsset{dbbmin/.style={color=dbbcolor,dashed,line width=1.0pt}}
\pgfplotsset{dbcmin/.style={color=dbccolor,dashed,line width=1.0pt}}
\pgfplotsset{dbdmin/.style={color=dbdcolor,dashed,line width=1.0pt}}

\pgfplotsset{dbamax/.style={color=dbacolor,solid,line width=1.0pt,mark=x ,mark options={solid}}}
\pgfplotsset{dbbmax/.style={color=dbbcolor,solid,line width=1.0pt,mark=o ,mark options={solid}}}
\pgfplotsset{dbcmax/.style={color=dbccolor,solid,line width=1.0pt,mark=square  ,mark options={solid}}}
\pgfplotsset{dbdmax/.style={color=dbdcolor,solid,line width=1.0pt,mark=triangle,mark options={solid}}}

\definecolor{LPMTrunc}{named}{TangoSkyBlue2}
\definecolor{LPMTruncInit}{named}{TangoSkyBlue2}
\definecolor{RFIR}{named}{TangoScarletRed3}
\definecolor{RFIRInit}{named}{TangoScarletRed1}
\definecolor{existing}{named}{TangoChameleon3}
\definecolor{G0Hat}{named}{TangoOrange2}
\definecolor{GVXI}{named}{G0Hat}
\definecolor{reference}{named}{black}

\definecolor{heuristic}{named}{TangoPlum2}
\definecolor{observed}{named}{TangoOrange3}

\definecolor{best}{named}{TangoAluminium6}

\definecolor{G0HatFill}{named}{TangoOrange1}
\definecolor{GVXIFill}{named}{G0HatFill}
\definecolor{LPMTruncFill}{named}{TangoSkyBlue1}
\definecolor{RFIRFill}{named}{TangoScarletRed1}
\definecolor{existingFill}{named}{TangoChameleon1}
\definecolor{bestFill}{named}{TangoAluminium3}

\definecolor{FRFMean}{named}{TangoPlum3}
\definecolor{FRFSingle}{named}{TangoPlum1}
\definecolor{FRFNoise}{named}{TangoChocolate1}

\pgfplotsset{FRFMean/.style={color=FRFMean,mark=*,mark options={solid},only marks,medsmallmarkers}}
\pgfplotsset{FRFSingle/.style={color=FRFSingle,mark=square*,mark options={solid},only marks,smallmarkers}}
\pgfplotsset{FRFNoise/.style={color=FRFNoise,mark=x,mark options={solid},only marks,smallmarkers}}

\pgfset{number format/1000 sep={\,}}

\pgfplotsset{bandwidth/.style={area style,fill=TangoButter1,draw=TangoButter2,fill opacity=0.5}}
\pgfplotsset{goodestimate/.style={color=TangoChameleon2,line join=round}}
\pgfplotsset{badestimate/.style={color=TangoScarletRed2,solid,line join=round}}
\pgfplotsset{exact/.style={color=black,dashed,line width=0.75pt,line join=round}}

\pgfplotsset{LPMTruncmark/.append style={mark=*,mark options={solid}}}
\pgfplotsset{RFIRmark/.append style={mark=square*,mark options={solid}}}
\pgfplotsset{existingmark/.append style={mark=triangle*,mark options={solid}}}
\pgfplotsset{existingInitmark/.append style={mark=triangle,mark options={solid}}}
\pgfplotsset{LPMTruncInitmark/.append style={mark=o,mark options={solid}}}
\pgfplotsset{RFIRInitmark/.append style={mark=square,mark options={solid}}}
\pgfplotsset{bestmark/.append style={mark=diamond*,mark options={solid}}}
\pgfplotsset{G0Hatmark/.append style={mark=asterisk,mark options={solid}}}

\pgfplotsset{GVXI/.style={color=GVXI,line width=1pt}}
\pgfplotsset{G0Hat/.style={color=G0Hat,line width=1.5pt}}
\pgfplotsset{existing/.style={color=existing}}
\pgfplotsset{LPMTruncInit/.style={color=LPMTruncInit,densely dashed}}
\pgfplotsset{LPMTrunc/.style={color=LPMTrunc}}
\pgfplotsset{RFIRInit/.style={color=RFIRInit,densely dashed}}
\pgfplotsset{RFIR/.style={color=RFIR}}
\pgfplotsset{best/.style={color=best}}

\pgfplotsset{smallmarkers/.append style={mark size=0.75pt}}
\pgfplotsset{medsmallmarkers/.append style={mark size=0.5pt}}
\pgfplotsset{tinymarkers/.append style={mark size=0.25pt}}
\pgfplotsset{extremelytinymarkers/.append style={mark size=0.05pt}}

\tikzset{annotation/.style={align=left,draw=black!0.2,font=\scriptsize,fill=white,fill opacity=0.8}}

%% TABLE STYLES

\pgfplotstableset{
	%zerofill=true,
	after row=[3pt],
	assign column name/.code={%
        \pgfkeyssetvalue{/pgfplots/table/column name}{\multicolumn{1}{c}{\multirow{2}{*}{#1}} }%
    },
	columns/method/.style={string type,column name={\shortstack{\textsc{Method}\\\phantom{0}}}}, 
	columns/contribution/.style={column name={\textsc{Contribution}}},
	columns/P000/.style={column type=r,column name={\shortstack[r]{\textsc{Min.}\\{$0\%$}}}},
	columns/P025/.style={column type=r,column name={\shortstack[r]{\vphantom{?}\\\textsc{$25\%$}}}},
	columns/P050/.style={column type=r,column name={\shortstack[r]{\textsc{Median}\\{$50\%$}}}},
	columns/P075/.style={column type=r,column name={\shortstack[r]{\vphantom{?}\\\textsc{$75\%$ }}}},
	columns/P100/.style={column type=r,column name={\shortstack[r]{\textsc{Max.}\\{$100\%$}}}},
	columns/contribution/.style={column name={\shortstack[r]{\textsc{Contribution}\\ \vphantom{0}}},
	postproc cell content/.append code={
            %\ifnum1=\pgfplotstablepartno
                \pgfkeysalso{@cell content/.add={}{\%}}%
            %\fi
        },
	},
	every head row/.style={before row=\toprule, after row={\\\midrule}},
    every last row/.style={after row=\bottomrule}
}