### 2015-10-29 front-and-back-matter/frontmatter.tex
Set up glossaries

\input{front-and-back-matter/summary.inc.tex}
\input{front-and-back-matter/acknowledgements.inc.tex}

\cleardoublepage
\addcontentsline{toc}{chapter}{Contents}
\tableofcontents
\clearpage

\KOMAoption{listof}{leveldown,totoc}

\listoffigures
\listoftables
% \TOCInsert{section}{List of Listings}   \lstlistoflistings %TODO: fix this

\printglossary[type=\acronymtype,style=long,toctitle={Acronyms}] %TODO: change format, set order, change title to "List of Abbreviations"
\glsaddall[types=symbols]
\printglossary[type=symbols,style=altlong4col,toctitle={List of Symbols}]

\newpage