### 2015-12-03 config/glossaries.inc.tex

\usepackage[sanitize={symbol=false},toc,acronym,section=section,nonumberlist,xindy]{glossaries}

\renewcommand*{\acronymname}{List of Abbreviations}
\renewcommand*{\glssymbolsgroupname}{List of Symbols}

\renewcommand{\newacronym}[4][]{\newglossaryentry{#2}{type=\acronymtype,
name={#3},
description={#4},
text={#3},
first={#4~(#3)},
plural={#3\glspluralsuffix},
firstplural={#4~(#3\glspluralsuffix)},
#1}}

\newglossary[slg]{symbols}{sls}{slo}{\glssymbolsgroupname}
%\newglossary{symbols}{sym}{sbl}{List of Symbols}
\newcommand{\newsymbolglos}[3]{\newglossaryentry{#1}{type=symbols,%
                                 name={\ensuremath{#2}},sort={#1}%
                               symbol={\ensuremath{#2}},description={#3}}}
\newcommand{\sym}[1]{\glssymbol{#1}}

%% TERMS

%% SYMBOLS
% \newsymbolglos{0-placeholder}{\placeholder}{Placeholder symbol}

\newsymbolglos{A-SetC}{\ComplexNumbers}{Set of the complex numbers}
\newsymbolglos{A-SetN}{\NaturalNumbers}{Set of the natural numbers}
\newsymbolglos{A-SetNnozero}{\NaturalNumbersWithoutZero}{Set of the natural numbers without the number $0$}
\newsymbolglos{A-SetRpos}{\PositiveRealNumbers}{Set of the positive real numbers}
\newsymbolglos{A-SetRneg}{\NegativeRealNumbers}{Set of negative real numbers}
\newsymbolglos{A-SetRnozero}{\RealNumbersWithoutZero}{Set of the positive real numbers without the number $0$, i.e. $\RealNumbers \without \set{0}$}
\newsymbolglos{A-SetQ}{\RationalNumbers}{Set of the rational numbers}
\newsymbolglos{A-SetR}{\RealNumbers}{Set of the real numbers}
\newsymbolglos{A-SetZ}{\IntegerNumbers}{Set of integer numbers}

\newsymbolglos{B-imagunit}{j}{The imaginary unit, $j = \sqrt{-1}$}
\newsymbolglos{B-epsilon}{\eps}{The numerical precision of a floating point number (\code{eps}), e.g. $\eps \approx 2.2 \cdot 10^{-16}$ for \code{double} precision}

\newsymbolglos{count-mc}{n_{\mathrm{MC}}}{Number of Monte Carlo runs}
\newsymbolglos{count-theta}{n_{\theta}}{Number of parameters}

\newsymbolglos{E-approx}{\approx}{is approximately equal to \ldots}
\newsymbolglos{E-bydef}{\isdef}{is defined as \ldots}
\newsymbolglos{E-propto}{\propto}{is proportional to \ldots}
\newsymbolglos{E-distributed}{\distributedAs}{is distributed as \ldots}

\newsymbolglos{lcm}{\lcm}{Least common multiple}

\newsymbolglos{matrix-eye}{\I_n}{ $n \times n$ identity matrix}

\newsymbolglos{norm-hinf}{\norm[\infty]{\placeholder}}{\Hinf norm of the system $\placeholder$}
\newsymbolglos{norm-two}{\norm[2]{\placeholder}}{$\Hardy_{2}$ norm of the system $\placeholder$}
\newsymbolglos{norm}{\norm[p]{\placeholder}}{$p$-norm of a signal or system $\placeholder$}

\newsymbolglos{operator-cardinality}{\numberOf \placeholder}{Number of elements (cardinality) in the set \placeholder}
\newsymbolglos{operator-exp}{\expval{\placeholder}}{Expected value of \placeholder}
\newsymbolglos{operator-ln}{\ln \placeholder}{Natural logarithm of \placeholder}
\newsymbolglos{operator-conj}{\conj \placeholder}{Complex conjugate of \placeholder}
\newsymbolglos{operator-log}{\log_{B} \placeholder}{Logarithm in base $B$ of \placeholder{}  (when $B$ is omitted, $B=10$ is implied)}
\newsymbolglos{operator-matrix-ht}{\placeholder^{\HT}}{Hermitian (complex conjugated) transpose of the matrix \placeholder}
\newsymbolglos{operator-matrix-pinv}{\pinv{\placeholder}}{Moore-Penrose pseudo-inverse of the matrix \placeholder}
\newsymbolglos{operator-matrix-tran}{\placeholder^{\TT}}{Transpose of the matrix \placeholder}
\newsymbolglos{operator-round-ceil}{\ceil{\placeholder}}{\placeholder{} rounded upwards (\code{ceil})}
\newsymbolglos{operator-round-floor}{\floor{\placeholder}}{\placeholder{} rounded downwards (\code{floor})}
\newsymbolglos{operator-round-round}{\round{\placeholder}}{\placeholder{} rounded towards the nearest integer, as in \glsentrytext{IEEE} 754-2008 (\code{round})}

\newsymbolglos{sys-damping}{\damping}{Relative damping of a system}
\newsymbolglos{sys-resonance}{\wn}{Natural (resonance) frequency of a system}
\newsymbolglos{sys-timeconst}{\tau}{Time constant of a system}

\newsymbolglos{w}{\omega}{Angular frequency (pulsation)}

%% ACRONYMS
\newacronym{BLA}{BLA}{best linear approximation}
\newacronym{BTLS}{BTLS}{bootstrapped total least squares}

\newacronym{CRLB}{CRLB}{Cramér-Rao lower bound}
\newacronym{CZT}{CZT}{chirp $Z$-transform}

\newacronym{DC}{DC}{direct current}
\newacronym{DCkernel}{DC kernel}{Diagonal Correlated (DC) kernel}
\newacronym{DFT}{DFT}{discrete Fourier transform}
\newacronym{DSA}{DSA}{dynamic signal analyzer}
\newacronym{DTFT}{DFT}{discrete-time Fourier transform}
\newacronym{DUT}{DUT}{device under test}

\newacronym{ETFE}{ETFE}{empirical transfer function estimate}

\newacronym{FFT}{FFT}{fast Fourier transform}
\newacronym{FRF}{FRF}{frequency response function}

\newacronym{GTLS}{GTLS}{generalized total least squares}

\newacronym{IDFT}{IDFT}{inverse discrete Fourier transform}
\newacronym{IEEE}{IEEE}{Institute of Electrical and Electronics Engineers}
\newacronym{IFAC}{IFAC}{International Federation of Automatic Control}
\newacronym{iid}{i.i.d.}{independent and identically distributed}

\newacronym{LML}{LML}{log marginal likelihood}
\newacronym{LPM}{LPM}{Local Polynomial Method}
\newacronym{LRIC}{LRIC}{Local Rational method with Iterative Cost function}
\newacronym{LRM}{LRM}{Local Rational Method}
\newacronym{LS}{LS}{least-squares}
\newacronym{LTI}{LTI}{linear time-invariant}

\newacronym{MIMO}{MIMO}{multiple-input multiple-output}
\newacronym{MISO}{MISO}{multiple-input single-output}
\newacronym{MLE}{MLE}{maximum-likelihood estimator}
\newacronym{MSE}{MSE}{mean squared error}

\newacronym{OE}{OE}{output-error}

\newacronym{PSD}{PSD}{power spectral density}

\newacronym{RARX}{RARX}{regularized auto-regressive exogenous}
\newacronym{RFIR}{RFIR}{regularized finite impulse response}
\newacronym{RMS}{RMS}{root-mean-square}
\newacronym{RMSE}{RMSE}{root mean squared error}

\newacronym{SIMO}{SIMO}{single-input multiple-output}
\newacronym{SISO}{SISO}{single-input single-output}
\newacronym{SPICE}{SPICE}{Simulation Program with Integrated Circuit Emphasis}
\newacronym{SNR}{SNR}{signal to noise ratio}

\newacronym{TRIMM}{TRIMM}{Transient Impulse Modeling}

\newacronym{VUB}{VUB}{Vrije Universiteit Brussel}

\makeglossaries