### 2016-02-16 front-and-back-matter/titlepages.tex
Update title page + summary

%\promotortitle{Promotor/Promotors}
\promotor{
  \jurymember{prof. dr. ir.}{Johan Schoukens}{Vrije Universiteit Brussel (VUB)}
  \jurymember{dr. ir.}{Tom Oomen}{Eindhoven University of Technology}
}

\title{Development of User-Friendly\\System Identification Techniques}

\author{ir. Egon Geerardyn}

\jury{
  \jurymember[president]{prof. dr. ir.}{Johan Deconinck}{VUB -- ETEC}
  \jurymember[vice-president]{prof. dr. ir.}{Patrick Guillaume}{VUB -- MECH}
  \jurymember[secretary]{dr. ir.}{Koen Tiels}{VUB -- ELEC}
  \jurymember{prof. dr. ir.}{Keith R. Godfrey}{University of Warwick}
  \jurymember{prof. dr. ir.}{Tomas McKelvey}{Chalmers University of Technology}
  \jurymember{dr. ir.}{Bart Peeters}{Siemens Industry Software -- LMS}
}

\date{\today~~{\color{gray}(\texttt{\VCRevision})}}

\maketitlepage