### 2016-02-28 ch04-lrmhinf/paper.tex

% \renewcommand{\glsgroupskip}{}
% \glsdisablehyper

\input{\thisDir/styles.inc.tex}

\newcommand{\wnValue}{\ensuremath{0.95 \unit{rad/s}}}
\newcommand{\dampingValue}{\ensuremath{0.025}}
\newcommand{\TsValue}{\ensuremath{1 \unit{s}}}

% \section{Introduction}
% \label{sec:Intro}

% Robustness is of key importance in feedback controlled systems, since feedback can lead to performance degradation or even closed-loop instability. 
% An important example includes an \gls{AVIS}, where feedback is used to isolate high-precision equipment from external disturbances. 
% The underlying feedback control principle is skyhook damping~\citep{Karnopp1995}. 
% However, the performance of skyhook damping is limited by high-frequent parasitic resonance phenomena. 
% Such model uncertainties can be taken into account explicitly in a robust control design~\citep{Zhang2005HybridAvis} for approaches based on \Hinf{} optimization.
% However, the uncertainty estimation in these references is based on rough prior assumptions and may be inaccurate. 
% On the one hand, this can lead to potentially dangerous results, since no stability and performance guarantees can be given if the estimated uncertainty is too small. 
% On the other hand, this can also lead to overly conservative controllers if the estimated uncertainty is too large.

% Several approaches have been developed to determine bounds on the model uncertainty. First, model validation techniques have been developed~\citep[see e.g.][]{Smith1992,Xu1999}.
% However, these results are typically overly optimistic~\citep{Oomen2009UncEstim}.
% Second, model error modeling approaches based on parametric system identification~\citep{Ljung1999MEM} with explicit characterization of bias and variance errors have been proposed, but these require a significant amount of user intervention.
% Third, non-parametric identification approaches have been adopted, e.g.~\citep{vandeWal2002,deVries1994}. 
% In~\citep{vandeWal2002}, an identified frequency response function is used directly to evaluate the \Hinf{} norm on a discrete frequency grid. 
% In~\citep{deVries1994}, an extended method is presented that bounds the error in between frequency points in a worst-case fashion. 
% However, such worst-case methods are well-known to be overly conservative~\citep[Section 9.5.2]{Vinnicombe2001}. 
% Fourth, recently a data-driven \Hinf{} norm estimation has been developed in~\citep{Wahlberg2010,Oomen2014ILH}.
% These methods rely on a sequence of iterative experiments and directly delivers an estimate of the \Hinf{} norm, and combines an optimal experiment design while essentially taking inter-grid errors into account.
% A recent application of this method in~\citep{Oomen2014ILH} has revealed that these iterative methods lead to significantly more accurate \Hinf{} estimates compared with traditional frequency response-based methods, thereby underlining the importance of the inter-grid error.
% Unfortunately, these methods require a sequence of dedicated iterative experiments.
% This increases the measurement time significantly, especially for \gls{MIMO} plants.
% When measurement time is expensive, this is a severe impediment.

% Although important steps have been made in model-error-modeling, existing methods involve a trade-off between accuracy, measurement time, and user intervention.
% The aim of the present paper is to develop a golden mean between traditional non-parametric techniques (third approach) and a full parametric model (second approach).
% The proposed $\Hinf{}$ norm estimation procedure exploits recent developments in frequency domain estimation the \gls{LPM}~\citep{Schoukens2009LPM} and \gls{LRM}~\citep{McKelvey2012LRM} as introduced in~\citep{Geerardyn2014IFAC,Geerardyn2014ISMA}.
% Before, these \gls{LPM} and \gls{LRM} techniques have been used to enhance at-grid frequency response estimates.
% In the present paper, these methods are extended to provide both an at-grid and inter-grid evaluation of the peak value, as is required for the $\Hinf{}$ norm. 
% The performance of these local parametric models with respect to robust control design is investigated, and bounds on the required measurement time to obtain a reliable estimate of the \Hinf{} norm are obtained.
% Preliminary results appear in~\citep{Geerardyn2014IFAC,Geerardyn2014ISMA}.
% The present paper presents the full procedure, including more theoretical treatment of the $\Hinf{}$ norm estimation procedure, and a description of the required measurement time.

% \vspace{-1em}
% \paragraph*{Outline}
% \label{par:toc}
% The problem is formulated in \secref{sec:Problem}.
% In \secref{sec:LPMHinf} the novel method to estimate the $\Hinf{}$ norm by means of the \gls{LRM} is introduced.
% In \secref{sec:simulation} the method is illustrated on simulations, and the required measurement time is determined.
% The technique is demonstrated on measurements of an \gls{AVIS} in \secref{sec:measurement}.
% Finally, the results and some challenges are discussed in \secref{sec:conclusion}.

% \glsresetall % reset the usage of all abbreviations s.t. intro and body get the explanation at first use

% \section{Problem Formulation}
% \label{sec:Problem}
% Robust control based on \Hinf{} optimization requires a nominal parametric model $\estimated\Plant$ and a bound on the model error $\Delta$. 
% After the model $\estimated\Plant$ is determined, either through first principles modeling or system identification, it remains to determine a bound on the model error $\Delta$.
% The \Hinf{} norm of the error system $\Delta$, i.e.
% \begin{equation}
%   \gamma 
%      = 
%        \infnorm{\Delta} 
%     \isdef 
%        \sup_{\omega} 
%          \abs{\Delta(\omega)}
%   \label{eq:HinfNormSISO}
% \end{equation}
% for a stable \gls{SISO} system, or its weighted form $\infnorm{W \Delta V}$, is a measure that serves an important purpose in many robust control design methodologies that specifies the `size' of the model class for which a controller is to be designed~\citep{Skogestad2005}.

% To estimate $\gamma$, only $\Delta(\omega_{\star})$ is of interest, where
% $ \omega_{\star} 
%      \isdef 
%        \arg
%          \sup_{\omega} 
%            \abs{\Delta(\omega)}
%   \text{.}
%   \label{eq:omega-star}
% $
% However, in a finite measurement time, non-parametric approaches only allow access to a finite and discrete grid of frequencies $\omega_k \in \DFTGrid$.
% As such, \eqref{eq:HinfNormSISO} can be rewritten as
% \begin{equation}
%   \gamma \isdef \sup_{\omega} \abs{\Delta(\omega)}
%          = \sup 
%              \set{
%                 \ezbrace{\sup_{\omega \in \DFTGrid  } \abs{\Delta(\omega)}}{\gamma_{\mathrm{FRF}}}
%                 \;,\;
%                 \ezbrace{
%                 \sup_{\omega \in \InterGrid} \abs{\Delta(\omega)}}{\gamma_{\mathrm{IG}}}}
%   \label{eq:gamma-split-grid}
% \end{equation}
% where $\Omega \isdef \openInterval{0}{\frac{2\pi}{\Ts}}$ is the measured frequency range for a sampling time $\Ts$.
% Two cases can be distinguished:
% \begin{enumerate}
%   \item $\omega_{\star} \in \DFTGrid$ such that $\abs{\Delta(\omega_{\star})} = \gamma =\gamma_{\mathrm{FRF}}$ is observed directly on the discrete frequency grid,
%   \item $\omega_{\star} \not\in \DFTGrid$, i.e. the extremal value $\abs{\Delta(\omega_{\star})}  = \gamma = \gamma_{\mathrm{IG}}$ is determined by the inter-grid behavior.
% \end{enumerate}
% The use of $\gamma_{\mathrm{FRF}}$ is standard in \gls{FRF}-based approaches, e.g.~\citep{vandeWal2002}.
% Its expected value over different realizations of the input signal $\expectedValue{\gamma_{\mathrm{FRF}}} \leq \gamma$ due to the limited frequency resolution.
% Such an underestimate may lead to overly optimistic controller designs.

% The main contribution of this paper is to  estimate the inter-grid term $\gamma_{\mathrm{IG}}$ without the need for a (global) parametric model.
% Instead, a  realistic value of $\gamma$ is obtained using parametric models that are valid over small frequency ranges.

% \subsection{Set-up}
% \TODO{overgang $\Delta(q)$ naar output-error}

% Consider an output-error \gls{LTI} discrete-time \gls{SISO} system $\Delta(q)$ excited by an input signal $u_{\Delta}(n)$.
% For an infinitely long data record, this can be described in the time-domain as
% \begin{equation}
%   y_{\Delta}(n) 
%     = 
%       y_{\Delta0}
%       + v(n) 
%     = 
%       \Delta(q)u_{\Delta}(n) 
%       + H(q)e(n)
%   \label{eq:sysDeltaInfData}
% \end{equation}
% where $q^{-1}$ is the lag operator, $e(n)$ is white Gaussian noise with unity variance, and both $\Delta(q)$ and $H(q)$ are stable causal real-rational functions.
% For a limited data record ($n \in \set{0,\ldots,N-1}$), the transient contributions $t_{\Delta}(n)$ lead to:
% \begin{equation}
%     y_{\Delta}(n) 
%       = 
%         \Delta(q)u(n) 
%         + H(q)e(n) 
%         + t_{\Delta}(n)
%     \text{.}
%     \label{eq:sysDeltaFiniteData}
% \end{equation}
% By applying the \gls{DFT}
% \begin{equation}
%   X(k) 
%        = 
%          \frac{1}
%               {\sqrt{N}} 
%          \sum_{n=0}^{N-1} 
%             x(n) \exp \left( \frac{- j 2 \pi k n }{N} \right)
%          \quad   
%          \forall k \in \mathbb{Z}
%   \label{eq:DFT}
% \end{equation} 
% to both sides of~\eqref{eq:sysDeltaFiniteData}, one obtains:
% \begin{equation}
%   Y_{\Delta}(k) 
%                 = 
%                   \Delta(\omega_k) U_{\Delta}(k) 
%                   + T_{\Delta}(\omega_k) 
%                   + V(k) 
%   \text{.}
%   \label{eq:sysDeltaFiniteDataFD}
% \end{equation}
% The index $k$ corresponds to the $k^{\text{th}}$ frequency bin with frequency $\omega_{k} = 2\pi k / (N \Ts)$ such that the frequency grid $\DFTGrid \isdef \set{\left.\omega_k \right| k = 1,\ldots,N-1}$.

% The presented results in this paper directly extend to closed-loop systems, as is shown in \secref{sec:control}.
% The results in this paper are presented in discrete time, but the extension to the continuous time case is straightforward.

% \section{\Hinf{} Norm Estimation Procedure}
% \label{sec:LPMHinf}
% To estimate $\gamma_{\mathrm{IG}}$, we propose to identify local parametric models in the frequency domain.
% The frequency axis is partitioned into continuous segments $\Omega_k \isdef \openInterval{ \omega_{k-1} }{ \omega_{k+1} }$ in view of the local validity of the local models.
% In particular, the inter-grid aspect of the $\Hinf$ norm is estimated as
% \begin{equation}
%   \gamma_{\mathrm{IG}} = 
%     \sup \set{ \left. 
%                 \sup_{\omega \in \Omega_k} \abs{\LocalModel{\Delta}_k(\omega)}
%                 \right|
%                 k = 1,\ldots,N-1
%                 }
%    \label{eq:gamma-inter-grid}
% \end{equation}
% where the $\LocalModel{\Delta}_{k}$ is the local model around the ($k$th) frequency $\omega_k$.
% An example of such local models is shown in \figref{fig:interpol-quantities}.
% Two alternative methods allow to evaluate~\eqref{eq:gamma-inter-grid} practically.

% \begin{figure}
%   \centering
%   \setlength{\figurewidth}{0.8\columnwidth}
%   \setlength{\figureheight}{0.68\figurewidth}
%   \input{\thisDir/figs/interpolation-show.tikz}
%   \caption{Illustration of the local models $\LocalModel{\Delta}_k$ around each frequency bin $\omega_k$ for $k\in\set{1,\ldots,6}$.
%   The bottom part shows the equivalence of the frequency segments $\Omega_k$ and the intervals $\set{\omega \given{k_L(\omega) = k \vee k_R(\omega) = k}}$.}
% \label{fig:interpol-quantities}
% \end{figure}

% \vspace{-3em}
% \paragraph*{Direct method}
% % \subsubsection{Direct method}
% One can implement the definition~\eqref{eq:gamma-inter-grid} of $\gamma_{\mathrm{IG}}$ directly.
% Hence, it boils down to solving the optimization problems
% \begin{equation}
%   % \gamma_{\mathrm{IG}}(k) 
%   % = 
%       \max_{\omega \in \Omega_k} 
%       \abs{\LocalModel{\Delta}_k(\omega)}
%       \label{eq:gamma-direct}
% \end{equation}
% for $k \in \set{1,\ldots,N-1}$.
% These are tractable since $\omega$ is one-dimensional and restricted to the intervals $\Omega_k$.
% Unfortunately, this approach offers no insight in the behavior of $\abs{\Delta(\omega)}$, e.g. around the peak amplitude.

% \vspace{-2em}
% \paragraph*{Interpolation method}
% To allow inspection of $\abs{\Delta(\omega)}$ in~\eqref{eq:HinfNormSISO} for arbitrary values of $\omega \in \Omega$, we propose an interpolation approach to estimate $\gamma$ and to allow visual inspection of $\abs{\Delta(\omega)}$.
% It can be understood intuitively that due to the limited validity domain of the local models, an estimate of $\Delta(\omega)$ can only be obtained from $\LocalModel{\Delta}_k(\omega)$ when $\abs{\omega_k-\omega}$ is small.
% To this end, two functions that indicate respectively the left and the right adjacent local models for frequency $\omega \in \Omega$ are defined:
% \begin{align}
%   k_L(\omega) & 
%                 \isdef
%                   \max
%                   \set{k 
%                     \given{
%                       \omega_k \in \DFTGrid, 
%                       \omega_k \leq \omega}
%                   }\\
%   k_R(\omega) & 
%                 \isdef
%                   \min
%                   \set{k 
%                     \given{
%                       \omega_k \in \DFTGrid, 
%                       \omega_k \geq \omega}
%                   }
%   \text{.}
%   \label{eq:kL-kR-indexing-functions}
% \end{align}
% Using the $k_L(\omega)$ and $k_R(\omega)$ functions, two candidate values for $\abs{\Delta(\omega)}$ are obtained: $\abs{\LocalModel{\Delta}_{k_L(\omega)}(\omega)}$ and $\abs{\LocalModel{\Delta}_{k_R(\omega)}(\omega)}$.
% In view of obtaining the maximum amplitude, these are `interpolated' to obtain a single estimate:
% \begin{equation}
%   \abs{\Delta_{\mathrm{LRM}}(\omega)}
%     \isdef
%     \sup \set{
%        \abs{\LocalModel{\Delta}_{k_L(\omega)}(\omega)},
%         \abs{\LocalModel{\Delta}_{k_R(\omega)}(\omega)}
%     }
%     \label{eq:DeltaLRMInterpolation}
% \end{equation}
% by retaining the largest both.
% It follows that
% \begin{align}
%   \widehat{\norm[]{\Delta_{\mathrm{LRM}}}}_{\infty} 
%     & \isdef
%     \sup_{\omega\in\Omega}
%     \abs{\Delta_{\mathrm{LRM}}(\omega)} \\
%    &  =
%       \sup_{\omega\in\Omega} \; 
%       \sup \set{
%                  \abs{\LocalModel{\Delta}_{k_L(\omega)}(\omega)}, 
%                  \abs{\LocalModel{\Delta}_{k_R(\omega)}(\omega)}}
%   \label{eq:gammaInterpolOptim}
%   \text{.}
% \end{align}

% \vspace{-2em}
% \paragraph*{Remarks:}
% The optimization problems~\eqref{eq:gammaInterpolOptim} and~\eqref{eq:gamma-direct} can, e.g. be solved by griding the frequency $\omega$.
% The peak amplitudes obtained using~\eqref{eq:gammaInterpolOptim} and~\eqref{eq:gamma-direct} are identical since the local models $\LocalModel{\Delta}_k$ are evaluated over identical frequency ranges
% $% \begin{equation}
%   \Omega_k = \set{\omega \given{k_L(\omega) = k \vee k_R(\omega) = k}}
% $ % \end{equation}
% for both methods (but in different order) and aggregated using the supremum operator, which is indifferent to the order of its arguments.
% In \figref{fig:interpol-quantities} the functions $k_L(\omega)$ and $k_R(\omega)$ are visualized and the equivalence with $\Omega_k$ is apparent.

% \subsection{The \glsentrydesc{LRM}}
% \label{sec:LRM}
% The local models $\LocalModel{\Delta}_k$ are estimated using the \gls{LRM}~\citep{McKelvey2012LRM}.
% The \gls{LRM} exploits the fact that the transfer function $\Delta(\omega)$ and the leakage term $T_{\Delta}(\omega)$  in \eqref{eq:sysDeltaFiniteDataFD} are smooth functions of the frequency.
% As such, $\Delta$ and $T_{\Delta}$ are approximated locally by rational functions around each bin $k$:
% \begin{align}
%   \Delta(\omega_{k+r})     
%     &\approx 
%         \frac{\sum_{i=0}^{N_{B}} b_{i}(k) r^{i}}{1 + \sum_{i=1}^{N_{A}} a_{i}(k) r^{i}} 
%       = 
%         \frac{B_k(r)}{A_k(r)} 
%       \isdef 
%         \LocalModel{\Delta}_{k}(\omega_{k+r})
%       \label{eq:LRMTaylorSystem} \\
%   T_{\Delta}(\omega_{k+r}) 
%     &\approx 
%         \frac{\sum_{i=0}^{N_{T}} t_{i}(k) r^{i}}{1 + \sum_{i=1}^{N_{A}} a_{i}(k) r^{i}}  
%       = 
%         \frac{T_k'(r)}{A_k(r)} 
%       \isdef 
%         \LocalModel{T}_{k}(\omega_{k+r})
%   \label{eq:LRMTaylorTransient}
%   \text{.}
% \end{align}
% Its local parameters $a_{i}(k)$, $b_{i}(k)$ and $t_{i}(k)$ are estimated by considering~\eqref{eq:sysDeltaFiniteDataFD} in the local window $\LocalWindow = \set{-N_W,\ldots,+N_W}$ around each bin $k$.
% This yields
% \begin{equation}
%   Y_{\Delta}(k+r) 
%                  \approx
%                     \Delta(\omega_{k+r}) U_{\Delta}(k+r) 
%                     + T_{\Delta}(\omega_{k+r}) 
%     \quad
%     \forall r \in \LocalWindow
%     \text{.}
%     \label{eq:LocalWindowSpectra}
% \end{equation}
% By substituting the local models $\LocalModel{\Delta}_k$ and $\LocalModel{T}_k$, one obtains
% \begin{align}
%   Y_{\Delta}(k+r) 
%                   & \approx
%                     \LocalModel{\Delta}_k(\omega_{k+r}) U_{\Delta}(k+r) 
%                     + \LocalModel{T}_k(\omega_{k+r}) 
%                     & \forall r \in \LocalWindow \\
%                   & = 
%                     \frac{B_k(r)}{A_k(r)} U_{\Delta}(k+r) 
%                     + 
%                     \frac{T'_k(r)}{A_k(r)} 
%                     & \forall r \in \LocalWindow
%     \text{.}
%     \label{eq:LocalWindowSpectraLRM}
% \end{align}
% To make this expression linear in the parameters $a_i(k)$, both sides are multiplied by $A_k(r)$ as in~\citep{Levy1959}.
% By doing so, the local linear-least-squares cost function
% \begin{equation}
%   \sum_{r=-N_W}^{N_W} 
%   \left| 
%     A_k(r) Y_{\Delta}(k+r) - B_k(r) U_{\Delta}(k+r) - T'_{k}(r)
%   \right|^2
%   \label{eq:localCostLRM}
% \end{equation}
%  around each bin $k$ can be constructed.
% We denote $\lrm{N_W,N_B,N_A,N_T}$ to be the \gls{LRM} with the orders $N_B$, $N_A$, $N_T$ and bandwidth $N_W$ defined above.
% For $N_A=0$ the \gls{LRM} simplifies to the \gls{LPM} which we shall denote as $\lpm{N_W,N_B,N_T}$.

% For a more detailed overview of  the \gls{LPM} and \gls{LRM} see~\citep{Schoukens2009LPM,Pintelon2010LPM1,Pintelon2010LPM2,McKelvey2012LRM}.

% \section{Simulation}
% \label{sec:simulation}
% To allow the generalization of the results of the simulations, we first introduce a way to decompose complex systems into simpler systems.
% The behavior of these simpler systems then allows to predict the behavior for these more complex systems.

% Any real-rational system $\Delta$ can be decomposed in a parallel structure of lower-order sub-systems.
% As a result, systems can often be well-represented, in a small frequency range, by a sum of a few low-order sub-systems (often first or second order for e.g. mechanical systems)~\citep[Section 2.2.2]{Gawronski2004}.
% To estimate $\norm{\Delta}$ mainly the resonant sub-systems are of interest.
% Assuming such resonances are sufficiently separated in the frequency,~\citep{Schoukens2013LPMerror,Geerardyn2013TIM,Gawronski2004} show that results for a single resonance are applicable to more complex systems.
% This is done by considering only the dominant resonance -- \emph{viz.} the one that is the most relevant for the $\Hinf{}$ controller design.
% By normalizing the conclusions for the time constant $\tau$ (or the $3\unit{dB}$ bandwidth), the results are generally valid for different values of the damping $\xi$.
% Specifically, the system's $\tau \isdef (\xi\omega_{\mathrm{n}})^{-1}$ dictates the required measurement time~\citep{Schoukens2013LPMerror}, where $\xi$ is the relative damping and $\omega_{\mathrm{n}}$ the undamped natural frequency.

% \subsection{A concrete simulation}
% Consider a discrete time system $\Delta$ (see \figref{fig:exampleIntergrid}) as unmodeled dynamics in an output-error setting:
% \begin{equation}
%   \Delta(z) 
%     =  \frac{0.45373 z + 0.44752}
%             {z^2 - 1.0595 z + 0.96079}
%   \label{eq:O2sysDT}
%   \text{,}
% \end{equation}
% which has $\tau = 50 \unit{samples}$ (i.e. $\xi=0.02$ and $\omega_{\mathrm{n}} = 1 \unit{rad/S}$).
% The input signal $u_{\Delta}$ is white Gaussian noise with unit variance and the disturbing noise variance $\sigma_v^2$ is chosen such that a \gls{SNR} 
% \begin{equation}
%   \SNR = \frac{\sqrt{\inv{N}\sum_n y_{\Delta0}^2(n) }}%
%               {\sqrt{\inv{N}\sum_n v^2(n) }}
%        = \frac{\sigma_{y_{\Delta0}}}{\sigma_v}
%        \approx \frac{\norm[2]{\Delta} \sigma_{u_{\Delta}}}{\sigma_v}
%   \label{eq:SNR}
% \end{equation}
% of $10$ is obtained.

% Classical \gls{SA} techniques~\citep{Bendat1980} can estimate the \gls{FRF} of $\Delta$.
% This is done by first splitting the input/output signals into $N_S$ segments of ${N}/{N_S}$ samples, windowing the segments to reduce leakage and applying the \gls{DFT}.
% By dividing the resulting output/input spectra, the \gls{FRF} is obtained.
% In this paper a Hann window and $N_S=1$ are used.
% Note that $N_S>1$  increases the measurement time by a factor $N_S$ to attain the same frequency resolution (but would improve  transient and noise rejection).

% \figref{fig:exampleIntergrid} illustrates for $N=97$ that $\gamma_{\mathrm{FRF}}$ for the \gls{SA}, $\lpm{5,4,4}$ and $\lrm{5,2,2,2}$ approaches yield unreliable estimates of $\infnorm{\Delta}$, since the frequency resolution is too coarse.
% The specific orders of  \gls{LRM}/\gls{LPM} have been chosen to  keep $N_W$ small while not restricting ourselves to linear models ($N_B=1$). 
% Slightly different orders (e.g. $\lrm{4,1,1,1}$) yield comparable results.

% \begin{figure}
%   \centering
%   \setlength{\figurewidth}{0.75\columnwidth}
%   \setlength{\figureheight}{0.68\figurewidth}
%   \input{\thisDir/figs/example-both.tikz}
%   %\input{\thisDir/figs/example-LRM-SNR1.tikz}
%   \caption{Simulation example revealing that $\gamma_{\mathrm{FRF}}$ (left) for both the \glsfirst{SA}~\legref{leg:example-SA} and \gls{LRM}~\legref{leg:example-LRM} underestimate the true $\gamma = \norm{\Delta} = 24$~\legref{leg:example-true-Hinf} due to limited frequency resolution.
%   Using the proposed \gls{LRM}-based approach, both the at-grid~($\gamma_{\mathrm{FRF}}$) and intergrid~($\gamma_{\mathrm{IG}}$) estimate are improved.
%   The interpolated \gls{LRM}~\legref{leg:example-LRM-interpol} and true $\Delta$~\legref{leg:example-true} almost coincide.
%   Hence, $\widehat{\norm[]{\Delta_{\mathrm{LRM}}}}_{\infty}$~\legref{leg:example-LRM-interpol-Hinf} approaches the theoretical $\gamma=24$.}
%   \label{fig:exampleIntergrid}
% \end{figure}

% It can be observed for $\gamma_{\mathrm{IG}}$ in \figref{fig:exampleIntergrid} that the \gls{LRM} is able to model the resonance well and yields a useful estimate $\infnorm{\Delta_{\mathrm{LRM}}} = 24$ in the simulation.
% Intuitively, the \gls{LRM} locally approximates the resonant pole which allows for a reliable inter-grid estimate.
% The \gls{LPM}, on the other hand, does not provide a reliable estimate.

% \subsection{Study of the experiment length}
% Frequency resolution, and hence the experiment length, is key to obtain a reliable \Hinf{} norm estimate.
% In this section, we determine the minimal experiment length needed for the proposed approach by means of simulations.

% We simulate the system \eqref{eq:O2sysDT} for $16\,100$ samples and discard the first $100$ samples to randomize the initial conditions.
% To mimic the effect of progressively measuring longer data records, the first remaining $N$ samples are used to estimate $\infnorm{\Delta}$ as in \secref{sec:LPMHinf}.
% The values of $N \in \closedInterval{35}{16 000}$ are chosen such that $\expectedValue{\gamma_{\mathrm{FRF}}(N)}$ is a non-decreasing function of $N$.
% These simulations are then repeated in a Monte Carlo simulation of $n_{\mathrm{MC}}=200$ runs where each run has a different realization of  $u_{\Delta}(n)$ and $v(n)$.
% The resulting sample average and standard deviation $\sigma$ are shown in \figref{fig:MC-run-length}.

% \begin{figure}
%   \centering
%   \setlength{\figurewidth}{0.75\columnwidth}
%   \setlength{\figureheight}{0.68\figurewidth}
%   \input{\thisDir/figs/MC-runlength.tikz}
%   %\input{figures/runLength.tikz}
%   \caption{The $\Hinf{}$ norm~\legref{leg:runlength-true} is underestimated significantly for short experiments ($N<20\tau$) by $\expectedValue{\gamma_{\mathrm{FRF}}}$~\legref{leg:runlength-frf} and the \gls{LPM}~\legref{leg:runlength-LPM}.
%   However, the proposed $\widehat{\norm[]{\Delta_{\mathrm{LRM}}}}_{\infty}$~\legref{leg:runlength-LRM} is already reliable from much shorter experiments ($N \geq 4 \tau$). A single observation for $N=97$~\legref{leg:runlength-special-case} is shown in \figref{fig:exampleIntergrid}. The shaded areas show the $\pm \sigma$ intervals which are mostly due to the noise.
%   }
% \label{fig:MC-run-length}
% \end{figure}

% It can be seen in \figref{fig:MC-run-length} that $N \geq 20\tau$ is required to attain a frequency grid that is dense enough to obtain a reliable $\gamma_{\mathrm{FRF}}$, i.e. $\expectedValue{\gamma_{\mathrm{FRF}}} \approx \gamma$.
% Starting from about $N=4 \text{ to } 5\tau$, the interpolated \gls{LRM} starts to yield good estimates for $\gamma_{\mathrm{IG}}$.
% For long experiments $(N \to \infty)$, the true $\norm{\Delta}$ is recovered.
% Practically, this implies that the interpolated \gls{LRM} beats the on-grid estimate using a data record that is only a fraction of the length.

% \section{Application to an \glsentryname{AVIS}}
% \label{sec:measurement}
% The proposed method is illustrated on the \gls{AVIS} shown in \figref{fig:avis}.
% The \gls{AVIS} consists of a support fixed to the ground and a payload platform, suspended on airmounts, that is free to move in six degrees of freedom.
% Six geophones measure the payload velocity and eight linear motors allow to actively compensate for vibration of the platform.
% To facilitate the presentation, only the vertical translation, i.e. a \gls{SISO} system,  is considered.

% \subsection{Measurement \& control framework}
% \label{sec:control}
% The experiments on the \gls{AVIS} are carried out in closed-loop (\figref{fig:tpc}) with $r_2 = 0$.
% The controller $\Controller=\experimental\Controller$  stabilizes the \gls{AVIS} but yields only moderate vibration isolation.
% The $r_1$ input consists of five periods of a random phase multisine~\citep{Geerardyn2013TIM} of $65\,536$ samples with excited bins such that $\omega_k \approx 1.001 \omega_{k-1}$  and $\Ts=1\unit{ms}$.

% \begin{figure}
%  \centering
%  \input{\thisDir/figs/tpcrobust.tikz}
%  \caption{The considered feedback configuration $\ClosedLoop{\ModelSet[],\Controller}$.}
%  \label{fig:tpc}
% \end{figure}

% \begin{figure}
%  \centering
%  \input{\thisDir/figs/dualYoula.tikz}
%  \caption{Dual-Youla parametrization of the closed-loop set-up  with controller $\Controller$,  model set $\ModelSet[]$ and nominal plant model $\estimated\Plant$.}
% \label{fig:dualYoula}
% \end{figure}

% The parametrization of the plant model set $\ModelSet[]$ conforms to the dual-\YK{} framework~\citep{Hansen1989,Anderson1998} that parametrizes $\ModelSet[]$ in terms of  all plants $P$ that are stabilized by the controller $C$:
% \begin{equation}
%   \ModelSet[] 
%     \isdef 
%       \set{
%           \left. 
%             \frac{\hat{N} + D_c \Delta}
%                  {\hat{D} - N_c \Delta} 
%           \right| 
%             \hinfnorm{\Delta} \leq \gamma
%         }
%   \label{eq:dualYoulaModelSet}
%   \text{,}
% \end{equation}
% where $C=N_c D_c^{-1}$ and $\estimated\Plant = \estimated{N}\estimated{D}^{-1}$ are decomposed as \gls{RCF} (see \figref{fig:dualYoula}).
% To estimate $\Delta$ and $\hinfnorm{\Delta}$, the signals $u_{\Delta}$ and $y_{\Delta}$ in~\figref{fig:dualYoula} are required.
% These can be computed from $u$ and $y$ directly~\citep{Anderson1998}.
% In addition, $u_{\Delta}$ is noise free under standard assumptions~\citep{Hansen1989}, hence the method of \secref{sec:LPMHinf} applies.

% The non-uniqueness of the \gls{RCF} can be exploited to satisfy additional conditions.
% In view of the robust control criterion
% \begin{equation}
%   \Crit(\Plant,\Controller) 
%     \isdef 
%       \infnorm{W \ClosedLoop{\Plant,\Controller} V}
%   \label{eq:controlCrit}
%   \text{,}
% \end{equation}
% the \glspl{RCF} are constructed such that
% $
%   \worstCase\Crit \left( \ModelSet[], \experimental\Controller \right)
%   \leq
%   \Crit \left( \estimated\Plant, \experimental\Controller \right)
%   + \gamma
% $
% where $\worstCase\Crit(\ModelSet[],C) \isdef \sup_{P\in\ModelSet[]} \Crit(P,C)$~\citep{Oomen2012SIRP}.
% % Obviously, $\Crit( \true\Plant, \experimental\Controller) \leq \worstCase\Crit ( \ModelSet[], \experimental\Controller )$ is guaranteed when $\true\Plant \in \ModelSet[]$.
% I.e. even for a weighted control criterion \eqref{eq:controlCrit}, there is no need to incorporate weights $W'$ and $V'$ into the estimation of the uncertainty bound as $\infnorm{W' \Delta V'}$.
% In this paper, an eighth-order $\estimated\Plant$~(\figref{fig:avis-frf}) is estimated.
% Some dynamics are left unmodeled and are hence part of $\Delta$~(\figref{fig:avisMeas}).

% \begin{figure}
%  \centering
%     \setlength{\figurewidth}{0.75\columnwidth}
%     \setlength{\figureheight}{0.68\figurewidth}
%     \input{\thisDir/figs/avis-frf.tikz}
%  \caption{Measured \gls{FRF} \legref{m2t:plantFRF} and estimated parametric model $\hat{P}$ \legref{m2t:plantModel} of the \glsfirst{AVIS}.}
%  \label{fig:avis-frf}
%  \label{fig:avis}
% \end{figure}

% \subsection{Results}
% \label{sec:resultsAvis}
% To estimate $\Delta(\omega_k)$ and $\hinfnorm{\Delta}$  using the proposed approach with $\lrm{6,2,2,2}$, an estimation dataset consisting of only a fourth of the excited bins is used.
% The full dataset that includes bins not used in the estimation is used to validate the interpolation results.

% The measured $\Delta$ is shown in \figref{fig:avisMeas} for the whole frequency band and \figref{fig:avisMeasZoom} for a frequency range containing $\omega_{\star}$.
% Two aspects of the proposed method are examined.
% First, the interpolation results for a continuous frequency range are validated.
% Second, the peak amplitudes of $\Delta(\omega)$ are inspected as these are  indicators for $\infnorm{\Delta}$.

% \begin{figure}
%  \centering
%     \setlength{\figurewidth}{0.75\columnwidth}
%     \setlength{\figureheight}{0.68\figurewidth}
%     \input{\thisDir/figs/avis-delta-full.tikz}
%  \caption{Measured (dual-Youla) $\Delta$ for the \gls{AVIS} using the \gls{LRM} and the proposed interpolation. The data in the rectangle is also shown in \figref{fig:avisMeasZoom}.}
% \label{fig:avisMeas}
% \end{figure}

% \begin{figure}
%  \centering
%     \setlength{\figurewidth}{0.75\columnwidth}
%     \setlength{\figureheight}{0.68\figurewidth}
%    \input{\thisDir/figs/avis4.tikz}
%  \caption{Measured $\Delta$ for the \gls{AVIS} using the \gls{LRM}~\legref{leg:avis4-LRMi} and the proposed interpolation. Clearly, the \gls{LRM} leads to a higher peak value~\legref{leg:avis4-LRM-hinf}, especially at $328\unit{Hz}$. The validity of the LRM local parametric model is confirmed by the validation measurement at the dense frequency grid~\legref{leg:avis4-validation}, which reveals excellent interpolation properties and a similar peak value~\legref{leg:avis4-valid-hinf}.}
% \label{fig:avisMeasZoom}
% \end{figure}

% \vspace{-2em}
% \paragraph*{Interpolation performance}
% \label{sec:avis-interpol}
% \figref{fig:avisMeas} and \figref{fig:avisMeasZoom} show that the estimation and validation data for \gls{LRM} are generally in agreement.
% In the frequency range shown in \figref{fig:avisMeasZoom}, the relative difference between both is at most $2\%$.

% \vspace{-2em}
% \paragraph*{Peak value estimates}
% \label{sec:avis-peak-value}
% Near $296\unit{Hz}$, a modest improvement of the peak estimate from $-15.3 \unit{dB}$ (at-grid) to $-14.0 \unit{dB}$ is obtained by the interpolation.
% The estimate near $\omega_{\star} \approx 328\unit{Hz}$ is improved more substantially.
% The proposed method \eqref{eq:gammaInterpolOptim} yields $\gamma_{\mathrm{IG}}=-2.5\unit{dB}$.
% $\gamma_{\mathrm{FRF}}\approx -9.9\unit{dB}$ for the estimation data, which means that a $7.4\unit{dB}$ improvement is achieved by the proposed method.
% This peak agrees well with the validation data, where the nearest (at-grid) value has an amplitude of $-3.2\unit{dB}$ (i.e. $8\%$ difference).

% This suggests that the local rational models with interpolation are a very good approximation to the actual $\Delta$ and also to estimate $\hinfnorm{\Delta}$ even if the actual peak does not coincide with the discrete frequency grid.

% \section{Conclusion and Further Research}
% \label{sec:conclusion}
% A reliable \Hinf{} norm estimation technique for high performance, non-conservative robust control design is investigated.
% In this paper, a new approach is presented that employs so-called local parametric \gls{LRM} models that lead to an enhanced estimate of $\norm{\Delta}$ of a \gls{SISO} error system by interpolating between neighboring local models.
% It is also shown that the \gls{LPM} is not a worthwhile alternative to the \gls{LRM} to estimate $\norm{\Delta}$.
% The technique is illustrated on a simulation example where the \gls{LRM}-based interpolation yields a reliable estimate of the $\Hinf{}$ norm.
% It is observed that the measurement time is reduced by almost an order of magnitude compared to techniques that only examine $\Delta(\omega)$ on the \gls{DFT} frequency grid.
% Using measurements on an \gls{AVIS}, it is illustrated that the interpolation-based results can substantially improve $\hinfnorm{\Delta}$ and/or reduce the measurement time four-fold.
% Furthermore, the results are validated by a detailed validation measurement.
% Ongoing research focuses on extending the technique to \gls{MIMO} systems and the study of its stochastic properties.