### 2016-02-28 config/blockdiag.inc.tex

\usetikzlibrary{positioning}  % positioning
\usetikzlibrary{decorations.pathmorphing}  % snake lines
\usetikzlibrary{matrix}                    % block alignment
\usetikzlibrary{arrows}                    % arrow heads
\usetikzlibrary{calc}                      % manipulation of coordinates
\usetikzlibrary{fit}

\pgfdeclarelayer{background}
\pgfdeclarelayer{foreground}
\pgfsetlayers{background,main,foreground}

%% Block diagram settings
\definecolor{weight}{named}{TangoButter3}
\definecolor{plant}{named}{TangoScarletRed2}
\definecolor{plantset}{named}{TangoOrange2}
\definecolor{modelset}{named}{plantset}
\definecolor{delta}{named}{black}
\definecolor{controller}{named}{TangoPlum2}

\definecolor{TPC}{named}{TangoAluminium4}

\tikzstyle{line}      = [thick]
\tikzstyle{block}     = [draw,rectangle,line,minimum height=2em,minimum width=2em]
\tikzstyle{weight}    = [block,color=weight]
\tikzstyle{controller}= [block,color=controller]
\tikzstyle{plant}     = [block,color=plant]
\tikzstyle{sum}       = [draw,circle,inner sep=0mm,minimum size=2mm]
\tikzstyle{connector} = [->,line]
\tikzstyle{branch}    = [circle,inner sep=0pt,minimum size=0.75mm,fill=black,draw=black]
\tikzset{groupbox/.style={draw,rectangle,line, fill opacity=0.2,fill=#1,color=#1}}
\tikzset{groupbox/.default=red}
\tikzstyle{weightbox} = [groupbox=weight]
\tikzstyle{none}      = [inner sep=0,outer sep=-0,fill=#1,#1]
\tikzset{none/.default=black}

\tikzset{groupcirc/.style={rectangle,fill opacity=0.2,fill=#1,color=#1}}
\tikzset{groupcirc/.default=red}