### 2016-03-03 front-and-back-matter/frontmatter.tex
Guidelines improved

\input{front-and-back-matter/acknowledgements.inc.tex}

\cleardoublepage
\addcontentsline{toc}{chapter}{Contents}
\tableofcontents
\clearpage

\KOMAoption{listof}{leveldown,totoc}

\listoffigures
\listoftables

%http://tex.stackexchange.com/questions/74857/toc-like-list-of-definitions-using-theorem-environments
\renewcommand{\listtheoremname}{List of Guidelines}
\listoftheorems[ignoreall,show=guideline]

% \TOCInsert{section}{List of Listings}   \lstlistoflistings %TODO: fix this

\printglossary[type=\acronymtype,style=long,toctitle={List of Abbreviations}, title={List of Abbreviations}] %TODO: change format

\glsaddall[types=symbols]
\printglossary[type=symbols,style=altlong4col,title={List of Symbols},toctitle={List of Symbols}]

\newpage