### 2016-03-31 ch04-lrmhinf/lrmhinf.tex

\chapter{Non-parametric resonance modeling}
\label{sec:hinf}
\def\thisDir{ch04-lrmhinf}
\tikzsetfigurename{ch04fig}
\myEpigraph{Design is not just what it looks like and feels like. Design is how it works.}{Steve Jobs}{}
\researchBasedOn{Geerardyn2014IFAC,Geerardyn2014ISMA}

\TODO{refer to book chapter of Tom}

\input{\thisDir/paper.tex}

\begin{subappendices}
  \input{\thisDir/ch04appendices.inc.tex}
\end{subappendices}