### 2016-04-04 config/mechanical.inc.tex
Add figures about vibration isolation

\usetikzlibrary{calc}
\usetikzlibrary{patterns}
\usetikzlibrary{decorations.pathmorphing}
\usetikzlibrary{decorations.markings}
\usepackage{contour}
% http://tex.stackexchange.com/questions/18472/tikz-halo-around-text
% http://tex.stackexchange.com/questions/19348/how-to-draw-a-line-passing-through-a-point-and-perpendicular-to-another
% http://tex.stackexchange.com/questions/13933/drawing-mechanical-systems-in-latex

\tikzstyle{spring}=[thick,decorate,decoration={coil,pre length=0.5em,post length=0.5em,segment length=0.5em,amplitude=0.75em}] %try decoration=zigzag
\tikzstyle{damper}=[thick,decoration={markings,  
  mark connection node=dmp,
  mark=at position 0.5 with 
  {
    \node (dmp) [thick,inner sep=0pt,transform shape,rotate=-90,minimum width=15pt,minimum height=3pt,draw=none] {};
    \draw [thick] ($(dmp.north east)+(2pt,0)$) -- (dmp.south east) -- (dmp.south west) -- ($(dmp.north west)+(2pt,0)$);
    \draw [thick] ($(dmp.north)+(0,-5pt)$) -- ($(dmp.north)+(0,5pt)$);
  }
}, decorate]

\tikzstyle{force}=[everyMechanical,-latex,thick]
\tikzstyle{sensorLine}=[everyMechanical,<->,very thick]

\colorlet{gndcolor}{TangoAluminium5}
\colorlet{skycolor}{TangoSkyBlue2}
\colorlet{darkSkycolor}{TangoSkyBlue3}
\colorlet{masscolor}{TangoAluminium1}

\tikzstyle{everyMechanical}=[draw,outer sep=0pt,thick,font=\bfseries]

\tikzstyle{mass}=[everyMechanical,minimum width=3em,minimum height=1.5em,draw=black,fill=masscolor]
\tikzstyle{mechController}=[mass,very thick,fill=white,minimum height=2.5em,minimum width=2em]

\tikzstyle{ground}=[everyMechanical,pattern color=gndcolor,pattern=north east lines,draw=none,minimum width=3em,minimum height=1.5em]
\tikzstyle{sky}=[ground, pattern color=skycolor,font={\color{darkSkycolor}}]

\tikzstyle{groundline}=[everyMechanical,color=gndcolor,line width=1pt]
\tikzstyle{skyline}=[groundline,color=skycolor]