### 2016-04-19 ch03-lrm/proofLOOCV.tex

\begin{subappendices}
  \section{Proof of the \glsentrytext{LOOCV} statistic for linear models}

  We follow an approach similar to \citet{Hyndman2014LOOCV} and \citet[Section 12.3.2]{Seber2003} to prove that for a linear model, one can easily compute the \gls{PRESS}/\gls{LOOCV} statistic by means of the so-called `hat-diagonals', the diagonal entries of the hat matrix $H$.
  This is relevant, e.g. for local linear models used in the \gls{LRM} and hence also the \gls{LPM}.

  We consider the linear model
  \begin{equation}
    Z = K \theta + V
  \end{equation}
  with $Z \in \ComplexMatrix{N\times1}$, $\theta \in \ComplexMatrix{\nth \times 1}$, $K \in \ComplexMatrix{N \times \nth}$.
$V \in \ComplexMatrix{N \times 1}$ is a \gls{iid} complex gaussian random variable.
The least-squares solution of such a linear model is:
\begin{equation}
  \hat{\theta} = \left(K^{\HT} K \right)^{-1} K^{\HT} Z = \pinv{K} Z
\end{equation}
such that the estimated output $\hat{Z}$ is given by
\begin{equation}
  \hat{Z} = K \pinv{K} Z = H Z
\end{equation}
where $H$ is the so-called `hat matrix' that projects the measurements $Z$ onto the estimates $\hat{Z}$.

The \gls{PRESS} statistic used in \gls{LOOCV} is defined as~\citep[Chapter 12]{Seber2003}
\begin{equation}
  \PRESS \isdef 
    \frac{1}{N} 
      \sum_{i=1}^{N}
      \abs{
        Z_i - \ignoring{i}{\hat{Z}}
      }^2
\end{equation}
where $\hat{Z}_i$ is the $i^{\text{th}}$ row of $\hat{Z}$ and $\ignoring{i}{\hat{Z}}$ denotes the estimated $\hat{Z}_i$ when the $i^{\text{th}}$ data point (i.e. row) is removed from the estimation problem.
For linear models, this is equivalent to
\begin{equation}
  \PRESS_{\mathrm{linear}} = 
     \frac{1}{N}
     \sum_{i=1}^{N}
     \abs{
       \frac{Z_i - \hat{Z}_i}
                {1 - H_{ii}}
     }^2
     \text{.}
\end{equation}
Note that in this last expression, no terms $\ignoring{i}{\hat{Z}}$ occur, such that the statistic can be computed without having to compute $N$ additional linear models.
In this appendix, we prove that these last two expressions are equivalent for linear models.

\paragraph{Proof}
\TODO{proof}
Denote the design matrix $\ignoring{i}{K}$ to be the design matrix $K$ where the $i^{\text{th}}$ row has been removed:
\begin{equation}
  \ignoring{i}{K}
  \isdef
  \begin{bmatrix}
    K_{1,1} & K_{1,2} & \cdots & K_{1,\nth}\\
   && \vdots\\
    K_{i-1,1} & K_{i-1,1} & \cdots & K_{i-1,\nth}\\
    K_{i+1,1} & K_{i+1,1} & \cdots & K_{i+1,\nth}\\
    && \vdots\\
    K_{N,1} & K_{N,2} & \cdots & K_{N,\nth}\\
  \end{bmatrix}
\end{equation}

\begin{lemma}[Matrix Inversion Lemma]\label{lem:matrix-inversion-lemma}
The Woodbury formula, states that
\begin{equation*}
\left(A+UCV \right)^{-1} 
= 
A^{-1} - A^{-1} U \left(C^{-1}+VA^{-1}U \right)^{-1} VA^{-1}
\end{equation*}
when the dimensions of the matrices allow for the multiplications~\citep[Section 3.2.2]{matrixcookbook}.
\end{lemma}

\TODO{evt. Seber2003/ChristensenXXXX: QR decomposition to compute $H$}

\end{subappendices}