### 2016-04-21 front-and-back-matter/acknowledgements.inc.tex
Uniform commands: part 1

\chapter{Acknowledgements}
\myEpigraph{We're all going to die, all of us, what a circus! That alone should make us love each other but it doesn't. We are terrorized and flattened by trivialities, we are eaten up by nothing.}{Charles Bukowski}{The Captain is Out to Lunch and \\ the Sailors have taken over the Ship}
% \myEpigraph{In the End, we will remember not the words of our enemies, but the silence of our friends.}{Martin Luther King Jr.}{}
%FIXME m2t team
%FIXME ELEC
%FIXME proffen, technici, secretaressen
%FIXME anna/maarten
%FIXME office mates
%FIXME EIT people
%FIXME jury
%FIXME christian larsson, istvan, MST
%FIXME mikaya, John
%FIXME CST
%FIXME Rick, Annemiek, Matthijs, Frank
%FIXME Johan
%FIXME Tom
%FIXME family
%FIXME family-in-law
%FIXME alexandra
%FIXME pixie