### 2016-07-22 ch03-lrm/leakage.tex
Stub: intro leakage

\section{Transients and Leakage in the Frequency Domain}
\label{app:nparam:leakage}

\disclaimer{
This appendix is based on multiple expositions regarding transient contributions in the frequency domain.
For a more in-depth discussion, we refer to \citep{Pintelon1997ARB}, \citep{Pintelon1997Transient}, \citep{Schoukens1999}, \citep[Section 2]{McKelvey2012LRM}, and \citep[Sections 2.6.3 and 6.3.2]{Pintelon2012}.
}

\TODO{derive transient contributions}

\TODO{illustrate the smoothness of leakage}

\TODO{make figures}