### 2016-08-02 cover/coverfigure.tex
Add Cover Figure

% arara: clean: { files: [coverfigure.auxlock, coverfigure.log] }
% arara: lualatex: { shell : yes, action : nonstopmode, options : "-8bit -file-line-error -halt-on-error" }
% arara: clean: { files: [coverfigure.aux, coverfigure.out] }

\RequirePackage{luatex85} %http://tex.stackexchange.com/questions/315025/lualatex-texlive-2016-standalone-undefined-control-sequence
\documentclass[crop=true,tikz=true,border=0.5cm]{standalone}

\usepackage{amsmath}
\input{config/definitions.tex}
\input{config/graphics.inc.tex}
\tikzexternaldisable
% \input{config/styles.inc.tex}

% \pagecolor{TangoSkyBlue3}

\colorlet{LRM}{TangoScarletRed2}
\colorlet{LRMemph}{TangoScarletRed3}
\colorlet{validation}{TangoChameleon1}

\colorlet{validation}{white}

\tikzset{line cap=round}
\tikzset{line join=round}
\tikzset{miter limit=20}

\pgfplotsset{frf/.style={color=frf,solid,mark options={solid},mark size=2pt}}
\pgfplotsset{interpol/.style={line width=3pt,solid,mark options={solid}}}

\pgfplotsset{LRMgrid/.style={color=LRM,mark size=4pt,only marks,mark=*,mark options={solid,color=LRMemph}}}

\pgfplotsset{validation/.style={color=validation,only marks,mark=square*,mark size=4pt,line width=2pt,mark options={solid,line width=2.5pt}}}

\pgfplotsset{clip mode=individual,}

\def\thisDir{ch04-lrmhinf}
% \newlength\figureheight
% \newlength\figurewidth
\setlength{\figurewidth}{36.2cm}
\setlength{\figureheight}{20cm}

\begin{document}
  \input{avis4.tikz}
  % \newpage
  % \input{avis-delta-full.tikz}
\end{document}